### 2022-08-22 short-article/short-article.tex
update short article and appendix

% !TEX program = XeLaTeX
\documentclass[12pt,ansiapaper]{article}

@@

% Dots in toc
\renewcommand{\cftsecleader}{\cftdotfill{\cftdotsep}}

@@

% Packages I've been using for different reasons
\usepackage{hyperref}
\usepackage{dcolumn}
\usepackage{graphicx}
\usepackage{float}
\floatplacement{figure}{H}
\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows}
\usetikzlibrary{positioning}
\usepackage{mathtools}
\usepackage{caption}

% English
\usepackage[english]{babel}
\usepackage[english]{isodate}
\cleanlookdateon

@@

% Title
\title{Vigilantism and Institutions:\\ Understanding Attitudes toward Lynching in Brazil\footnote{We thank Rosario Aguilar, Louise Araújo, Nicholas Barnes, Hannah Baron, Dara Kay Cohen, Nicholas Cowen, Daniel J. D'Amico, Miriam Golden, Umberto Mignozzetti, Eduardo Moncada, Irfan Nooruddin, Jonathan Obert, Brian J. Phillips, Catarina Roman, Alexandra Scacco, Livia Schubiger, Natán Skigin, Nicholas Rush Smith, and Georg Vanberg for their valuable comments. We received helpful feedback in seminars at EGAP, the Extra-legal Governance Institute at the University of Oxford, the Comparative Workshop at Brown University, the Kroc-Kellogg Peace, Conflict, Crime and Violence Workshop at Notre Dame, and the Severyns Ravenholt Seminar in Comparative Politics at the University of Washington. This study received IRB approval from Brown University in October 2020 (Protocol 2009002803). Replication materials are available at \url{https://github.com/danilofreire/lynching-experiment-brazil}. We thank the Centre for the Study of Governance \& Society at King's College London and the Templeton Foundation for financial support.}}

@@

% Abstract
\begin{abstract}
\doublespacing \noindent Why do people support extrajudicial violence? In a series of survey experiments with respondents in Brazil, we examine how people view different motivations for lynchings, which characteristics of lynching scenarios garner greater support for lynching, and whether providing different types of information about lynching reduces support for it. We find that people do not believe that criminal justice institutions should be relied on to solve serious crimes and do believe that in such instances, community members have the right to take vengeance. In particular, our analysis finds that people strongly support the use of extrajudicial violence by families of victims against men who sexually assault and murder women and children. 
\vspace{.25cm}

@@

The regulation of violence is one of the most important functions of the state. When states control violence effectively, markets, politics, and civil society can flourish. However, in many places, violence is poorly regulated and nonstate armed actors---ranging from residents to rebel groups---use violence at their own discretion. The extralegal use of violence is also a clear violation of the rule of law \citep[48-49]{blair2020peacekeeping}. More generally, the challenge of monopolizing violence is at the heart of the Weberian definition of the state, and it is arguably one of the most fundamental questions of political economy. While political scientists have written extensively on civil wars, rebel governance, and criminal violence \citep{arjona2016rebelocracy, trejo2021high, barnes2017criminal}, they have written far less about vigilante violence. In this paper, we use survey experiments to better understand people's attitudes toward the use of extralegal violence.

Although lynchings and mob violence more generally have decreased in advanced democracies, it remains widespread in many parts of the world. Based on US State Department reports, \citet[33]{jung2020lynching} find that between 1976 and 2018, lynchings occurred in all regions of the world and in more than a hundred countries. In particular, Latin America has been severely affected by a recent increase in lynchings. Such a surge in citizen-led violence has reportedly been motivated by growing numbers of drug-related crimes and by police ineffectiveness \citep{mallen2014vigilantes}. Brazil provides a telling example. From 2011 to 2015, the country registered about 2,500 lynching episodes. In 2015, 173 people were killed by angry mobs---nearly one execution every two days \citep{barbara2015vigilantes, oliveira2016mob}. According to José de Souza Martins \citeyearpar{martins2015linchamentos}, who has studied lynchings in Brazil for more than thirty years, these figures are not only the highest in the country's history, but among the highest in the world. The people who participate in lynchings are typically  young men, but they also sometimes include teenage women and girls, elderly women, and even members of the local police \citep{moura2017linchamentos}.\footnote{Brazilian police forces in São Paulo kill hundreds of people every year in a way that looks similar to vigilantism \citep{willis2015killing}. Recent work also documents how organized crime groups in Brazil use extralegal violence in the process of governing favelas  \citep{magaloni2020killing}.}

There is no strong consensus about what factors explain people's participation in vigilantism. Some of the most common explanations focus on the role of state-based institutions. People living in places with weak or failed states might engage in vigilante activity to control crime and to enforce social order \citep{bancroft1887works}. However, others argue that it is growth in state capacity that motivates people to take justice into their own hands out of fear that the state will not punish perpetrators or will do so too lightly \citep{smith2019contradictions}. In Latin America in particular, vigilantism has been found to be associated with previous exposure to violence \citep{garcia2019anger}, untrustworthy law enforcement \citep{zizumbo2017community}, perceived insecurity \citep{ceobanu2011crime, godoy2004justice}, and economic and social inequality \citep{phillips2017inequality,godoy2006popular,arias2010violent}.

Part of the reason for the proliferation of findings is that studying vigilantism empirically presents a substantial challenge. Many studies focus on vigilantism along the American frontier in the mid-nineteenth century, drawing on archival and historical documents  \citep{brown1975strain, courtwright2009violent, obert2018six}. However, much relevant evidence needed to test competing theories has been lost to history. Qualitative and ethnographic research provides deep insights into the motivations and emotions surrounding social disorder, violence, and vigilantism \citep{godoy2002lynchings}. Yet, like all research approaches, this method has shortcomings, including recall bias and social desirability bias. Quantitative approaches that rely on observational data are likewise often limited in important ways, including that relevant data may be absent, data may imperfectly map onto theoretical concepts, and data may be collected and aggregated at a level of analysis far removed from the studied phenomenon. Finally, while a credible identification strategy is not necessary for making causal claims, it can often solve issues that arise from poor-quality data. Unfortunately, credible identification strategies are often not available.  As \citet[17]{bateson2020politics} notes, ``Field experiments on vigilantism would raise serious ethical concerns, and natural experiments are rare.''

In this paper, we use survey experiments to better understand individual perceptions of vigilantism. Based on the existing literature, we designed three survey experiments to understand lynchings in Brazil in particular. This empirical approach offers several distinct advantages. First, the data we collected come from the same level of analysis as the phenomenon itself: local residents. Second, online surveys are less susceptible to social desirability bias, and survey experiments help to elicit sensitive information---an especially important benefit in researching this topic \citep{grimm2010social}. Third, the experimental nature of the approach allows us to estimate the size of effects. Finally, we can also control for confounding, which is a potential concern with more qualitative methods.

One obvious downside to survey experiments is the difficulty in making inferences about the external validity of findings. \citet[17-18]{bateson2020politics} suggests that the closer that a fictional survey prompt matches the real world, the better the external validity. Accordingly, our first experiment is based on an actual account of a lynching in Jundia\'{i}, Brazil. In the second experiment, we include a wide range of lynching-scenario characteristics that allow respondents to consider detailed, realistic examples. Our third experiment examines whether providing different types of information can affect respondents' perceptions of lynchings. Our strategy here is similar to that of public marketing campaigns that governments often use to change public opinion. In this experiment, we provide information akin to government messaging to the general public.

In the first experiment, we attempt to understand how the underlying motivation for a lynching affects support for it. In particular, we test whether police absence, the slowness of the legal process, and the belief that punishment will not be harsh enough affect support. We find no effect for any treatment. Instead, combined with existing data on measures of police ineffectiveness and state capacity, we argue that the evidence is consistent with the claim that respondents simply do not view the police as a genuine source of assistance for serious crimes. Our second experiment is a conjoint experiment designed to see which characteristics of a lynching lead people to believe it is more or less justified. People believe that the victims of crimes or their family members have the right to seek revenge, especially for heinous or otherwise serious crimes. People see crimes such as forcible rape and murder as especially suitable for extrajudicial violence. In the final experiment, we provide information to respondents to attempt to reduce support for lynching. We remind respondents that lynching is a crime, that all citizens have human rights, and that lynchings can prompt violent vendettas. We find large negative effects of reminding respondents that lynching is a crime and that lynchings carry the risk of vendettas.

In studying lynching, this  paper contributes to several debates. First, we provide causal estimates of why Brazilians justify lynching, how vigilantes' motivation affects the justification, and how to reduce support for lynchings. Second, we provide evidence that is consistent with the common argument that vigilantism is driven by state absence or weakness. However, state absence or weakness alone is not enough to explain support for vigilantism. Like \cite{jung2020lynching}, we argue that people's beliefs and cultural norms are crucial to whether vigilantism is perceived to be justified. It is not simply the fact that the state is unavailable to enforce the law that leads people to support lynchings. Instead, it is the belief that it is not the proper role of the police to do so and that the response should instead be carried out by the family of the victim. Third, our findings inform the relationship between political development and informal norms. Even when legal institutions are present, norms  can undermine people's reliance on them. This is a major challenge in establishing the rule of law in developing countries, where residents often prefer informal solutions to crime over state-based solutions \citep[10]{blair2020peacekeeping}. As a result, the prevalence of lynchings undermines political development and state capacity, which can fall into a self-reinforcing cycle \citep{jung2020lynching}. As such, solving the problem of discretionary, extrajudicial use of violence is crucial for understanding political and economic development more generally.

\section{Literature Review and Hypotheses}
\label{sec:theory}

In an important recent article, Regina Bateson \citeyearpar{bateson2020politics} conceptualizes vigilantism and defines it as ``the extralegal prevention, investigation, or punishment of offenses'' (3). She identifies five key characteristics of vigilantism, each of which varies along a continuum (11--14). First, vigilantism can be carried out by an individual or by a group of people. Second, vigilantism ranges from nonviolent interventions to extremely violent ones, including even the torture and murder of someone who is suspected of committing a crime. Third, it can occur in private or public, and some acts of vigilantism are intentionally carried out in elaborate, public displays. Fourth, vigilantism can arise spontaneously or it can be highly organized and institutionalized. Finally, vigilantism can operate to prevent crime in the first place or to respond to crime during or after its occurrence (what Bateson calls ``defensive'' and ``offensive'' vigilantism). In this paper, we focus on lynchings, a form of vigilantism that is carried out by groups, with violence, in public, spontaneously, and in response to crime being committed.\footnote{Different types of vigilantism might arise for different reasons. For example, highly organized and institutionalized ``vigilance committees'' might be caused by entirely different social forces than are spontaneous lynchings. Accordingly, we focus on a single, specific type of vigilantism---lynching---throughout the paper.}

The study of vigilantism has a long and multidisciplinary history. Both classic works and contemporary research draw from a variety of fields and methods, including anthropology, criminology, history, sociology, and, more recently, political science \citep{moncada2017varieties}. These studies rely on historical, archival, observational, and statistical empirical approaches. This work---which strongly informs our survey design---offers several theories to explain variation in the prevalence of vigilantism.\footnote{We are obviously not testing all existing theories of vigilantism. This is partly because of constraints facing survey-based research. For example,  recent work shows that economic inequality drives Mexican \textit{autodefensa} activity \citep{phillips2017inequality}. Likewise, there are  long-run historical determinants that explain variation in armed resistance \citep{Osorio2019mexico}. Surveys are not especially conducive to testing those types of variables and theories. Instead, we focus on theories based on state presence, perceptions of legitimacy, and cultural norms, which are well suited to survey research.}  First, many scholars argue that vigilantism arises in response to the inability of states to effectively control crime and enforce social order \citep{bancroft1887works}. Historical studies of California identify state absence or weakness as a major cause of the emergence of vigilance committees and miners' courts during the gold rush (1848--55) \citep{hunt1920committees,rohrbough1997days,couttenier2017wild}.\footnote{There is substantial debate about the effectiveness of these activities \citep{clay2003order}. More generally, on extralegal institutions along the nineteenth-century frontier, see \cite{murtazashvili2013political}.} While gold miners often pursued justice in loose, disorganized, and fleeting ways \citep{mcdowell2002commons,mcdowell2007criminal}, some of these vigilance committees were highly organized and semipermanent \citep{bancroft1887works}.\footnote{\citet{taniguchi2016dirty} argues that the committee's actions in 1856 were not a response to a need for order, but instead a land grab from people who acquired land rights when Mexico controlled the region.} Outside of the United States, many studies likewise find that a desire for security and safety drive vigilantism. In Ghana, for example, the lack of criminal justice resources and the presence of corruption among security actors have been blamed for a rise of vigilantism between 1990 and 2000 \citep[416]{adinkrah2005vigilante}. Similarly, in  Guatemala since 1996, lack of effective state-based crime control was partly responsible for the rise of vigilantism \citep{godoy2006popular}. This research suggests the first hypothesis:

\begin{itemize}
  \item \textbf{Hypothesis 1:} Support for vigilantism will increase if people believe they cannot rely on the state to control crime and enforce social order.
\end{itemize}

One challenge to the hypothesis in the context of the US frontier is that vigilante activity was also widespread during roughly the same time along the East Coast and in the South, where states were far more established than along the Western frontier. In response, one theory argues that vigilantism in these locations instead arose along ``social frontiers'' \citep{obert2018keeping}. These are places where social relationships and social standing were ambiguous among different racial, ethnic, and class-based groups. If the state was not strong or stable enough to eliminate these ambiguities, then vigilantism emerged as a way to create and entrench social and political identities. Black Americans, in particular, were lynched in shocking numbers and frequency across the South well into the twentieth century \citep{dray2003hands,seguin2019national}. Likewise, based on studies of executions of Muslim men in Bosnia and a lynching in Jim Crow--era Maryland, Lee Ann Fujii \citeyearpar{fujii2017talk} argues that lynchings serve as ``violent displays'' that send signals to observers in the community and entrench social hierarchies and political identities.

While the racial prejudice that drove violence in the United States is clear, vigilantism more generally need not be driven primarily by racial tensions. For example, \citet{martins2015linchamentos} writes that lynchings in Brazil are not mainly due to racial prejudice. Local accounts report no significant social difference between victims and perpetrators; nor are Black Brazilians particularly targeted by lynch mobs \citep{band2015rocinha, pearson2018latam, sinhoretto2009linchamentos}. These events might be driven by racism, but they do not reproduce the stark, racially disparate patterns that existed in the United States historically \citep[23]{martins2015linchamentos}. \citet[8]{jung2020lynching} likewise note that ``\textit{racist} social control'' is a unique property of American vigilantism. Brazil's patterns thus pose a puzzle for existing explanations for mob violence and stand in contrast to the American experience \citep{smaangs2016doing, wood2011lynching}.

Recent research has shown that the relationship between states and vigilantism is more complicated than some early studies assumed. Based on extensive interviews and fieldwork in post-apartheid South Africa, Nicholas Rush \citet{smith2019contradictions} finds that people participate in vigilantism because the state is seen to be \textit{too} strong. Residents perceive a strengthening of the state, fuller support for human rights, and a stronger rule of law. As a result of these perceived changes, crime victims and other community members feel that the state will take too long to arrest, charge, and convict a perpetrator. If the perpetrator is released before trial, they might harm others or intimidate witnesses. Finally, people in South Africa fear that any punishment eventually received will not be severe enough. Accordingly, many South Africans believe it is right for people to take justice into their own hands. As a result, achieving legal reform and a  closer adherence to the rule of law can actually sparks greater use of extrajudicial violence. This suggests a second hypothesis:

\begin{itemize}
  \item \textbf{Hypothesis 2:} Support for vigilantism will increase if people believe states provide too much protection for offenders' rights. 
  \end{itemize}

Studies of vigilantism in Africa likewise find that religion, norms, and gender are powerful forces shaping the emergence and operation of vigilante groups. For example, in Nigeria, an increase in the influence of neoliberalism (alongside other political changes) drove vigilantism \citep[5]{pratten2008politics}. In Ghana, nearly 20 percent of the lynchings between 1990 and 2000 happened because people believed that someone (usually an older woman) had used ``mystical powers'' to wound someone's genitals or to cause other health problems  \citep[414]{adinkrah2005vigilante}. More generally, gender norms play an important role in the use of extrajudicial violence. The family members of people who are victims of a serious crime, such as forcible rape, often believe it tarnishes the reputation of the victim and their family. As a result, they believe they have the right to punish the offender. This suggests a third hypothesis:

\begin{itemize}
  \item \textbf{Hypothesis 3:} Support for vigilantism will increase when victims are female and their family carries out retribution.  
  \end{itemize}

In our three experiments, we argue that respondents do not believe that state-based legal institutions are the primary way to respond to serious crime. We find that people's attitudes toward vigilantism are shaped strongly by the characteristics of the incident that gives rise to it. In particular, people strongly support lynching by the family of a victim of a sexual offense or murder. Finally, our third experiment shows that support for lynchings can be reduced by reminding people of the penalties for lynching and of the risk of sparking vendettas.

@@

\section{What Motivates Support for Lynching?}
\label{sub:exp02}
We invited respondents from all regions of Brazil to participate in our experiment. Qualtrics recruited 2,406 adult Brazilian citizens in 2020 between October 30 and December 22 to take part in our survey experiment.\footnote{Our ethics statement is available in Section E of the supplementary material.} We included ten questions on demographics and other information that might influence the results. These include the respondent's age, gender, ethnicity, level of education, monthly family income relative to the minimum wage, political orientation (left to right), support for the death penalty, whether they had been victimized in the last year, their trust in the police, and their trust in the judicial system. We used these measures to disentangle heterogeneous effects in the main responses.

In our first experiment, we examined how the stated motivation for a lynching affects respondents' views about whether it is justified. In particular, we assessed the impact of three factors that have been cited as major drivers of vigilante justice: (1) police ineffectiveness; (2) slow response by the criminal legal system; and (3) demand for harsher punishment for offenders. Belief in the ineffectiveness of policing appears to be a strong predictor of vigilantism \citep{cruz2019determinants, garcia2019anger}. If police cannot or will not stop crime, then people often believe they have a right to do so themselves. Related to this, within the academic literature, lack of trust in the criminal legal system often motivates vigilantism \citep{godoy2004justice, smith2019contradictions}. This is often because of long criminal proceedings, which cause significant anxiety for the victims. Finally, we examined whether respondents believe that the legal punishment given to offenders is severe enough. In particular, we wanted to assess whether respondents are motivated by a desire for harsher penalties in Brazil's criminal legal system.

The experiment consisted of three treatment conditions and one control group. Respondents read an excerpt of a news article describing a real lynching case in Jundia\'{i}, Brazil. Jundia\'{i} is a well-known metropolitan city of about four hundred thousand residents in the state of S\~{a}o Paulo, and we chose it because it is not a particularly distinctive, violent, or unusual city. We slightly edited the original text so that respondents had no prior knowledge of the crime.\footnote{The original article is available at \url{https://jr.jor.br/2020/05/01/homem-e-linchado-na-vila-progresso}. Access: August 2020.} Consistent with Bateson's \citeyearpar[3]{bateson2020politics} emphasis on vigilantism as including ``prevention, investigation, or punishment of offenses,'' the vignette for the control group explains that the lynching took place while a crime was occurring. We asked respondents to show their level of support for mob violence using a 0--100 slider, where 0 means no support and 100 means full support.\footnote{We chose not to describe an especially heinous crime, such as murder or sexual assault, because we wanted people to respond to the treatment rather than the crime itself. By focusing on a less serious crime, the treatment is more salient. As shown in the second experiment, there is strong support for lynching for more serious crimes. The baseline of support for lynching in this vignette is about 36 percent (on a scale of 0 to 100), so the crime discussed appears sufficiently serious to justify lynching to a significant number of people.} Respondents in each of the three treatments read the same piece, but with one additional sentence explaining the motivations behind the lynching. Treatment 1 (which is motivated by Hypothesis 1) aimed to test the state-absence explanation. Hypothesis 2 claims that greater protection of offenders' rights increases vigilantism. Treatments 2 and 3 tested two versions of that hypothesis, in the form of slower judicial processes and insufficiently harsh punishments. The vignettes are as follows:

\begin{itemize}
\tightlist
\item
  \emph{Control group}: A man was lynched last Friday in Jundiaí, São Paulo. According to the neighbors, he tried to break into a house but was immobilized and beaten by members of the community.\footnote{In Portuguese: ``Um homem foi linchado na última sexta-feira em Jundiaí, São Paulo. De acordo com vizinhos, ele tentou invadir uma residência mas foi imobilizado e agredido por membros da comunidade.''}
\end{itemize}

\begin{itemize}
\tightlist
\item
  \emph{Treatment 1 (Police ineffectiveness)}: A man was lynched last Friday in Jundiaí, São Paulo. According to the neighbors, he tried to break into a house but was immobilized and beaten by members of the community. \textbf{One of the residents who took part in the lynching said they had beaten the suspect because ``the police never patrol the area.''}\footnote{In Portuguese: ``Um homem foi linchado na última sexta-feira em Jundiaí, São Paulo. De acordo com vizinhos, ele tentou invadir uma residência mas foi imobilizado e agredido por membros da comunidade. \textbf{Um dos moradores envolvidos no linchamento disse que eles agrediram o suspeito porque `a polícia nunca patrulha o local.'''}}
\end{itemize}

\begin{itemize}
\tightlist
\item
  \emph{Treatment 2 (Slow criminal legal system)}: A man was lynched last Friday in Jundiaí, São Paulo. According to the neighbors, he tried to break into a house but was immobilized and beaten by members of the community. \textbf{One of the residents who took part in the lynching said they had beaten the suspect because ``the judicial system is too slow and the perpetrator is on the street until the case is heard.''}\footnote{In Portuguese: ``Um homem foi linchado na última sexta-feira em Jundiaí, São Paulo. De acordo com vizinhos, ele tentou invadir uma residência mas foi imobilizado e agredido por membros da comunidade. \textbf{Um dos moradores envolvidos no linchamento disse que eles agrediram o suspeito porque ``a justiça é muito lenta e os criminosos ficam soltos até o julgamento.'''}}
\end{itemize}

\begin{itemize}
\tightlist
\item
  \emph{Treatment 3 (Demand for harsher legal punishment)}: A man was lynched last Friday in Jundiaí, São Paulo. According to the neighbours, he tried to break into a house but was immobilised and beaten by members of the community. \textbf{One of the residents who took part in the lynching said they had beaten the suspect because ``the judicial punishment is not harsh enough.''}\footnote{In Portuguese: ``Um homem foi linchado na última sexta-feira em Jundiaí, São Paulo. De acordo com vizinhos, ele tentou invadir uma residência mas foi imobilizado e agredido por membros da comunidade. \textbf{Um dos moradores envolvidos no linchamento disse que eles agrediram o suspeito porque `a punição da justiça não é dura o suficiente.'''}}
\end{itemize}

Before each vignette, respondents read the following text:

\begin{itemize}
\tightlist
\item
  You will be shown a news article. Please read it carefully. After you read the article, we will ask you one question about it.\footnote{In Portuguese: ``Uma notícia será apresentada para você. Por favor, leia a notícia com atenção. Após você ler o artigo, faremos uma pergunta sobre ele.''}
\end{itemize}

After the vignette, respondents were presented with this question:

\begin{itemize}
\tightlist
\item
  Do you think that the lynching was justified? Please use the slider below to indicate your opinion. For disagreement, use 0 to 49; for agreement, use 51 to 100. Please use 50 if you neither agree nor disagree.\footnote{In Portuguese: ``Você acha que o linchamento foi correto? Por favor, use a barra abaixo para indicar sua opinião. Para discordar, use de 0 a 49; para concordar, use de 51 a 100. Por favor, use 50 para não concordar nem discordar.''}
\end{itemize}

\begin{table}[htbp] \centering
  \caption{Average treatment effects for experiment 1}
  \label{tab:exp01main}
\begin{tabular}{@{\extracolsep{3pt}}lD{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3} }
\\[-1.8ex]\hline \\[-1.8ex]
\\[-1.8ex] & \multicolumn{4}{c}{\textbf{Lynching Support}\vspace{.5cm}} \\
\\[-1.8ex] & \multicolumn{1}{c}{(1)} & \multicolumn{1}{c}{(2)} & \multicolumn{1}{c}{(3)} & \multicolumn{1}{c}{(4)}\\
\hline \\[-1.8ex]
 Police do not patrol area & 1.115 &  &  &  \\
  & (1.861) &  &  &  \\
  Justice too slow &  & -0.289 &  &  \\
  &  & (1.927) &  &  \\
  Punishment not harsh enough &  &  & 0.438 &  \\
  &  &  & (1.921) &  \\
  Combined treatments &  &  &  & 0.443 \\
  &  &  &  & (1.545) \\
  Constant & 36.300^{***} & 36.300^{***} & 36.300^{***} & 36.300^{***} \\
  & (1.332) & (1.332) & (1.332) & (1.332) \\
 N & \multicolumn{1}{c}{1,161} & \multicolumn{1}{c}{1,111} & \multicolumn{1}{c}{1,103} & \multicolumn{1}{c}{2,215} \\
\hline \\[-1.8ex]
\multicolumn{5}{l}{$^{*}$p $<$ .1; $^{**}$p $<$ .05; $^{***}$p $<$ .01} \\
\multicolumn{5}{l}{Robust standard errors in parentheses.} \\
\end{tabular}
\end{table}

We randomized the treatment and control conditions for the entire respondent pool. This randomization procedure was independent of that of the other experiments. We carried out our hypothesis tests with OLS. We compared the average score given by the control group with the average score given by respondents in each treatment condition.

The first piece of evidence that the experiment provides concerns how baseline levels of support for lynching vary according to individual characteristics. Table \ref{tab:exp01-baselines} shows the impact of gender, race, and ideology on the subjects' tolerance of lynchings. Our dependent variable ranges from 0 to 100. We observed that men are more likely to support lynchings when compared with women. This is in contrast to research in Uganda, Tanzania, and South Africa, where women support mob vigilantism more than men do because they believe that they are less likely to be subject to false accusations \citep{wilke2021gender}.\footnote{On the relationships between gender and violence, see \citet{kadera2018gendered} and related papers.} In line with \citet{martins2015linchamentos}, we found little evidence that support for lynchings is driven primarily by outright racial prejudice.\footnote{We recognize that race is socially-constructed and that racial categories can be codified in multiple ways. We adopt the racial classification used by the Brazilian Census Bureau (IBGE) as it is widely employed in social research \citep[e.g.][24]{lapop2018brazil} and intuitive to our respondents. Nevertheless, our findings are also consistent with research that uses different, and often more nuanced, conceptions of race \citep{monk2016consequences, schwartzman2020colour}. } Ideology is strongly correlated with support for lynching, with a negative coefficient for citizens who identify as being on the center left or left.\footnote{The results are very similar when we estimated the models using data from our third experiment. Please refer to Section D.2 in the supplementary material for the more information.}

\begin{table}[ht] \centering
  \caption{Determinants of baseline levels of lynching support}
  \label{tab:exp01-baselines}
\begin{tabular}{@{\extracolsep{3pt}}lD{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3} D{.}{.}{-3} }
\\[-1.8ex]\hline \\[-1.8ex]
\\[-1.8ex] & \multicolumn{4}{c}{\textbf{Lynching Support}\vspace{.5cm}} \\
\\[-1.8ex] & \multicolumn{1}{c}{(1)} & \multicolumn{1}{c}{(2)} & \multicolumn{1}{c}{(3)} & \multicolumn{1}{c}{(4)}\\
\hline \\[-1.8ex]
 Male & 5.124^{***} &  &  & 4.945^{**} \\
  (\textit{Reference: female}) & (1.828) &  &  & (2.165) \\ \\
  Asian &  & 1.119 &  & -6.050 \\
  &  & (4.646) &  & (7.288) \\
  Mixed race &  & 0.724 &  & -5.256 \\
  &  & (2.467) &  & (4.026) \\
  White &  & -1.283 &  & -9.877^{***} \\
  (\textit{Reference: Black}) &  & (2.288) &  & (3.672) \\ \\
  Left &  &  & -11.895^{***} & -10.108^{***} \\
  &  &  & (2.388) & (3.199) \\
  Center left &  &  & -14.707^{***} & -17.484^{***} \\
  &  &  & (2.673) & (3.575) \\
  Center right &  &  & -3.019 & -5.047 \\
  &  &  & (2.881) & (3.838) \\
  Right &  &  & 0.610 & 2.722 \\
  (\textit{Reference: center}) &  &  & (2.328) & (3.235) \\ \\
  Constant & 33.105^{***} & 37.122^{***} & 41.804^{***} & 45.738^{***} \\
  & (1.251) & (2.109) & (1.793) & (4.160) \\
 N & \multicolumn{1}{c}{1,142} & \multicolumn{1}{c}{2,186} & \multicolumn{1}{c}{1,632} & \multicolumn{1}{c}{838} \\
\hline \\[-1.8ex]
\multicolumn{5}{l}{$^{*}$p $<$ .1; $^{**}$p $<$ .05; $^{***}$p $<$ .01} \\
\multicolumn{5}{l}{Robust standard errors in parentheses.} \\
\end{tabular}
\end{table}

There were no significant effects on any of the treatments. One possible interpretation of the null findings is that state presence and activity have no influence on attitudes toward lynching. If that interpretation is correct, we should reject Hypotheses 1 and 2.

Given the extensive literature discussed above and the qualitative evidence collected in the next experiment (discussed in Section \ref{sec:exp01}), an alternative explanation is that the null results reflect the fact that respondents simply have little to no faith in the criminal legal system in general. Several pieces of evidence support this interpretation. First, many crimes in Brazil are never solved. For example, a recent statistic indicates that the police only solve about 10 percent of the homicides in Brazil, suggesting that even for serious crimes, the police provide little help \citep{pearson2018latam}. Second, Brazilian police forces themselves carry out a substantial number of unjustified killings of people from disadvantaged and minority communities, which undermines their legitimacy and the sense that they can be relied on \citep{willis2015killing}.\footnote{Recent work on Mexico shows that implementing rule of law reforms can reduce police violence \citep{magaloni2020institutionalized}.} In 2019, police killed more than 1,800 people in Rio de Janeiro alone. Third, the Brazilian penal code allows an accused person to appeal each decision several times, so it can take several years (and sometimes more than a decade) before a criminal case is closed \citep{sousa2005utilizaccao}. Given this, people do not believe that a perpetrator will be punished in a timely matter even if put on trial. Finally, along with the rise of Jair Bolsonaro---the far-right president who has publicly called for people to take the law into their own hands \citep{brant2021armas, uol2021bolsonaro}---survey evidence suggests that people have grown considerably more punitive \citep{datafolha2018penademorte}.

>>>>

\documentclass[12pt,a4paper]{article}

@@

@@

% Packages I've been using for differen reasons
\usepackage{hyperref}
\usepackage{dcolumn}
\usepackage{graphicx}
\usepackage{float}
\floatplacement{figure}{H}
\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows}
\usetikzlibrary{positioning}
\usepackage{mathtools}
\usepackage{caption}

% English
\usepackage[english]{babel}
\usepackage[english]{isodate}
% \cleanlookdateon

@@

% Title
\title{Vigilantism and Institutions:\\ Understanding Attitudes Toward Lynching in Brazil\footnote{We thank Rosario Aguilar, Louise Araújo, Nicholas Barnes, Hannah Baron, Dara Kay Cohen, Nicholas Cowen, Daniel J. D'Amico, Miriam Golden, Richard Locke, Umberto Mignozzetti, Eduardo Moncada, Irfan Nooruddin, Jonathan Obert, Brian J. Phillips, Catarina Roman, Alexandra Scacco, Livia Schubiger, Natán Skigin, Nicholas Rush Smith, Richard Snyder, and Georg Vanberg for their valuable comments. We received helpful feedback in seminars at EGAP, the Extra-legal Governance Institute at the University of Oxford, the Comparative Workshop at Brown University, the Kroc-Kellogg Peace, Conflict, Crime and Violence Workshop at Notre Dame, and the Severyns Ravenholt Seminar in Comparative Politics at the University of Washington. This study received IRB approval from Brown University in October 2020 (Protocol 2009002803). Replication materials are available at \url{https://github.com/danilofreire/lynching-experiment-brazil}. We thank the Centre for the Study of Governance \& Society at King's College London and the Templeton Foundation for financial support.}}

@@

% Abstract
\begin{abstract}
\doublespacing \noindent Why do people support extrajudicial violence? In two survey experiments with respondents in Brazil, we examine which characteristics of lynching scenarios garner greater support for lynching and whether providing different types of information about lynching reduces support for it. We find that people often do support community members to take vengeance. In particular, our analysis finds that people strongly support the use of extrajudicial violence by families of victims against men who sexually assault and murder women and children. We also find that criminal punishment and the threat of vendettas reduce support, but appeals to the human rights of victims have zero effect on support for lynchings. 
\vspace{.25cm}

@@

The regulation of violence is a crucial function of the state. When states
control violence effectively, markets, politics, and civil society can flourish
\citep{besley2011pillars, north2009violence}. However, states still struggle to
maintain the monopoly of force in much of the Global South. Scholars have
written extensively about how rebel organizations build support to challenge
government institutions, yet we know little about what motivates local
extrajudicial violence in the developing world \citep{bateson2020politics,
smith2000new}. This omission is significant not only because local violence is
pervasive in weak states, but also because it can be even more costly than war-
or terrorism-related violence \citep{blair2017predicting}.

One of the most serious forms of extrajudicial violence is lynching, which can
be defined as ``incidents of physical violence committed by large numbers of
private citizens against one or more individuals accused of having committed a
``criminal'' offense, whether or not this violence resulted in the death of the
victim(s)'' \citep[645]{godoy2004justice}. Although lynchings occur in more
than one hundred countries in all regions of the world
\citep[33]{jung2020lynching}, Latin America has been particularly affected by a
sharp increase in vigilante violence. In recent years, countries like Brazil
and Venezuela have experience nearly one execution every two days, and lynching
episodes have also been reported in Guatemala, Mexico, Peru, and other
countries in the region \citep{cruz2019determinants, godoy2004justice,
barbara2015vigilantes}. These alarming figures suggest that, despite being
dismissed by policymakers as irrational and senseless, vigilantism enjoys
considerable approval in Latin America \citep{berg2011globalizing,
goldstein2005flexible}. This raises a number of puzzles: under which conditions
do citizens perceive lynchings as a legitimate response to crime? Are some
groups seen as more ``deserving'' of lynchings that others? And how can we
reduce pro-vigilantism attitudes in young, unequal democracies?

In this study, we run two survey experiments and use primary qualitative data
to understand attitudes toward lynchings in Brazil. First, we employ a conjoint
experiment with eight attributes to evaluate in which circumstances citizens
perceive lynchings as a justifiable punishment for crime. We also included
open-ended questions for respondents to elaborate on their decision process.
Finally, we use an information provision experiment to test the efficacy of
three strategies to reduce individual support for lynchings.

Our empirical approach offers several distinct advantages. First, the data we
collected come from the same level of analysis as the phenomenon itself: local
residents. Second, online surveys are less susceptible to social desirability
bias, and survey experiments help to elicit sensitive information---an
especially important benefit in researching this topic \citep{grimm2010social}.
Third, the experimental nature of the approach allows us to precisely estimate
the size of effects and compare different theoretical explanations
simultaneously. Finally, we can also control for confounding, which is a
potential concern in previous research on vigilantism.

Our results show that people's attitudes toward lynching are shaped strongly by
the characteristics of the incident that gives rise to it. In particular,
respondents express support for lynchings perpetrated by the family of a victim
of a sexual offense or murder. Subgroup analyses show that our findings are
consistent across different segments of the population. In line with
\citet{godoy2006popular} and \citet{martins2015linchamentos}, we find no
experimental or qualitative evidence that lynching support is primarily driven
by the race of the criminal. Our results also indicate that popular support for
lynchings can be substantially reduced by reminding individuals of the
penalties for lynching and of the risk of sparking vendettas, but not by
emphasizing human rights protections. This suggests that campaigns focusing on
the risks to vigilantes' personal security can be effective correctives to
pro-lynching attitudes and may strengthen support for the rule of law.

@@
